---
title: "Exercises 02"
author: "Jan Hohenheim"
date: "3/10/2022"
output: html_notebook
---

## Setup
```{r whipe}
rm(list = ls())
```

```{r loading}
library(readr)
library(ggplot2)
library(ggthemes)
library(ggfortify)
library(qqplotr)
library(dplyr)
library(GGally)
library(glue)
library(tidyr)
library(cowplot)
library(latex2exp)
library(purrr)
```

```{r data}
financing_healthcare <- read_csv("../all_datasets_bio144/financing_healthcare.csv")
head(financing_healthcare)
summary(financing_healthcare)
glimpse(financing_healthcare)

theme <- 
  theme(text = element_text(family = "Helvetica Neue")) +
  theme_solarized_2(light = FALSE)
```

## Task
First make / synthesise some normally distributed residuals, by drawing some random numbers from the normal distribution. The function to do this in R is rnorm().

resids <- rnorm(30, 0, 1)

Then use ggplot to plot the distribution of these residuals. You know how to do this! Don't forget to load the package that ggplot is in -- you will get the error "Error: could not find function "ggplot"" if you forget to do this.

In this case, the distribution really should look normal, as we drew numbers from a normal distribution!

The QQ-plot is quite easy to make:

qqnorm(resids)

And we can put a line on where we expect the points to lie if the data is normally distributed:

qqline(resids)

```{r recap}
resids <- data.frame(resids = rnorm(30, 0, 1))
plot_hist <- resids |>
  ggplot(aes(x = resids)) +
  geom_histogram(bins=15, alpha = 0.5, fill = "salmon") +
  ggtitle("Histogram of normal sample") +
  theme

plot_qq_norm <- resids |>
  ggplot(aes(sample = resids)) +
  stat_qq_band(fill = "salmon", alpha = 0.5) +
  stat_qq_line() +
  stat_qq_point() +
  ggtitle("QQ plot of normal sample") +
  theme

plot_grid(plot_hist, plot_qq_norm, ncol = 2)
```

```{r plot_dists}
plot_distribution <- function(data, name) {
  resids <- data.frame(resids = data)
  plot_hist <- resids |>
    ggplot(aes(x = resids)) +
    geom_histogram(bins = 15, alpha = 0.5, fill = "salmon") +
    ggtitle(glue("Histogram of {name} sample")) +
    theme
  plot_qq <- resids |>
    ggplot(aes(sample = resids)) +
    stat_qq_band(distribution = name, fill = "salmon", alpha = 0.5) +
    stat_qq_line(distribution = name) +
    stat_qq_point(distribution = name) +
    ggtitle(glue("QQ plot of {name} sample")) +
    theme
  
  list(plot_hist, plot_qq)
}

n <- 100
args <- list(
  list(name = "norm", data = rnorm(n, 0, 1)),
  list(name = "unif", data = runif(n)),
  list(name = "beta", data = rbeta(n, 5,2)),
  list(name = "beta", data = rbeta(n, 2, 5)),
  list(name = "lnorm", data = rlnorm(n, 0, 1)),
  list(name = "pois", data = rpois(n, 10)),
  list(name = "nbinom", data = rnbinom(n, 10, 0.5))
  )
plot_distribution_from_args <- function(args) do.call(plot_distribution, args)
plots <- distributions |>
  lapply(plot_distribution_from_args) |>
  flatten()
plot_grid(plotlist = plots, ncol = 2, align = "hv")
```
```{r plot_dists_by_side}
plot_distribution <- function(data, name) {
  resids <- data.frame(resids = data)
  plot_hist <- resids |>
    ggplot(aes(x = resids)) +
    geom_histogram(bins = 15, alpha = 0.5, fill = "salmon") +
    ggtitle(glue("Histogram of {name} sample")) +
    theme
  plot_qq <- resids |>
    ggplot(aes(sample = resids)) +
    stat_qq_band(distribution = name, fill = "salmon", alpha = 0.5) +
    stat_qq_line(distribution = name) +
    stat_qq_point(distribution = name) +
    ggtitle(glue("QQ plot of {name} sample")) +
    theme
  
  plot_grid(plot_hist, plot_qq)
}

n <- 100
args <- list(
  list(name = "norm", data = rnorm(n, 0, 1)),
  list(name = "unif", data = runif(n)),
  list(name = "beta", data = rbeta(n, 5,2)),
  list(name = "beta", data = rbeta(n, 2, 5)),
  list(name = "lnorm", data = rlnorm(n, 0, 1)),
  list(name = "pois", data = rpois(n, 10)),
  list(name = "nbinom", data = rnbinom(n, 10, 0.5))
  )
plot_distribution_from_args <- function(args) do.call(plot_distribution, args)
distributions |>
  lapply(plot_distribution_from_args)
```
