---
title: "Exercises 01"
author: "Jan Hohenheim"
date: "3/10/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r loading}
rm(list = ls())
library(readr)
library(ggthemes)
library(ggfortify)
library(tidyverse)
library(GGally)
library(glue)
library(tidyr)
library(latex2exp)

theme <- 
  theme(text = element_text(family = "Helvetica Neue")) +
  theme_solarized()
```

## How many observations?
The caste variable shows whether an individual is a worker or a lazy. How many individuals are in the worker caste (1 point)?
How many are lazy (1 point)?

Hint: use the table function.
```{r table}
mole_rat <- read_csv("../all_datasets_bio144/MoleRatLayabouts.csv")
mole_rat$caste |>
  table()
```

## OK distributions?
How do the distributions of the Mass and Energy variables look?

Hint: because there are not so many data points, you will need to choose less than 30 bins in your histogram (30 is the default).

```{r ok}
p1 <- mole_rat |>
  ggplot(aes(x = Mass)) +
  geom_histogram(bins = 20) +
  theme
p2 <- mole_rat |>
  ggplot(aes(x = Energy)) +
  geom_histogram(bins = 20) +
  theme
plot_grid(p1, p2)
```

## Who is lazy?
Using ggplot (of course) make a graph that will help you answer the biological question... do lazy and worker individuals differ in the energy expenditure, after one accounts for any difference in size.  
For a given mass, do worker or lazy individuals have higher energy expenditure?
```{r lazy}
mole_rat |>
  ggplot(aes(x = Mass, y = Energy, colour = caste)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme
```

##  How many degrees of freedom?

If you fit model D, how many degrees of freedom for error would remain?

D: lm(log10_Energy ~ caste * log10_Mass, dd)

```{r df}
# intercept, caste (2 - 1), mass, caste:mass
nrow(mole_rat) - 1 - 1 - 1 - 1
```

##  What slope?
Fit the model with the interaction (this is sometimes called the "full" model) and answer the following questions:

Rounded to two significant figures, what is the estimated difference in slope between the two castes (1 point)?

What is the standard error (rounded to two sig. figs.) (1 point)?

```{r slope}
dd <- mole_rat |>
  mutate(log10_Energy = log10(Energy),
         log10_Mass = log10(Mass))
m1 <- lm(log10_Energy ~ caste * log10_Mass, dd)
m1 |> summary()
```

# What difference in R-squared?
Fit the model with both main effects but no interaction.

What is the difference between the r-squared (R2) of this model, and the one with the interaction (rounded to two sig. figs. and ignore sign)?
```{r diffr2}
m2 <- lm(log10_Energy ~ caste + log10_Mass, dd)
m2 |> summary()

summary(m2)$r.squared - summary(m1)$r.squared
```

## Exercise 1 Figure

Now comes probably the most challenging part of the practical. Making a figure that shows the best supported statistical model, i.e. the one without an interaction term.

It is challenging because you should not do this:

ggplot(dd, aes(x=log10_Mass, y=log10_Energy, colour=caste)) +
    geom_point() +
    geom_smooth(method="lm")

You can't do this, because the lines on the graph have different slopes, and the best supported model does not have different slopes.

Hint: you will need to use the expand.grid and predict functions. You saw this in action before, in the multiple regression lesson.

OK, I give you a good start...

Have a look at the new dataset this creates. Then use it with predict. Then add it to the plot, just like I did in the body fat video lectures.

If you get done with that, make the graph look beautiful, have appropriate axes labels, etc. Also, try to make it work in black and white.
```{r fig}
new_data <- expand.grid(
  log10_Mass = seq(min(dd$log10_Mass), max(dd$log10_Mass), length = 100),
  caste = unique(dd$caste))
new_data |> glimpse()

predicted_energy <- m2 |>
  predict(newdata = new_data, interval = "confidence")

predicted_mole_rat <- new_data |>
  cbind(predicted_energy)

predicted_mole_rat |>
  ggplot(aes(x = log10_Mass, y = fit, colour = caste)) +
  geom_line() +
  geom_point(data = dd, aes(y = log10_Energy)) +
  geom_smooth(aes(ymin = lwr, ymax = upr), stat = "identity") +
  labs(title = "Model prediction of energy expenditure for constant mass",
       x = TeX("$\\log_{10}$(mass in g)"),
       y = TeX("$\\log_{10}$(energy expenditure in kcal)")) +
  theme
```
