---
title: "Exercise 2"
author: "Jan Hohenheim"
date: "3/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data


```{r q0}
rm(list=ls())
library("readr")
library("ggplot2")
library("dplyr")
library("GGally")
library("tidyr")
library(ggthemes)

financing_healthcare <- read_csv("../all_datasets_bio144/financing_healthcare.csv")
head(financing_healthcare)
```

## Questions 1 and 2

How many countries and years are in the dataset?

Hint: try using the summarise function in dplyr, and ask for the length of the unique list of countries.


```{r q1}
financing_healthcare |>
  summarise(countries=n_distinct(country), years=n_distinct(year))
```

## Question 3
How many rows have a non "NA" value for both of the two variables we're most interested in: health_exp_total and child_mort?

Hint: you can ask if a value in a variable is NA with the function is.na(). Or if you use !is.na() then you get the ones that are not NA. E.g. !is.na(my_variable) asks which values of my_variable are not NA.

Hint: Use the ampersand symbol to ask two things.

Hint: use filter in the dplyr library.

```{r q3}
financing_healthcare |>
  select(health_exp_total, child_mort) |>
  filter(!(is.na(health_exp_total) | is.na(child_mort))) |>
  count()
```

## Question 4

There are lots of interesting things we could do with this dataset, and we'll get to some in future weeks. For now lets do two things.

First, let us have a go at getting statistics for groups of observations, and then making a graph to visualise those statistics.

The first thing to do is to reduce the dataset a bit more:

Make a dataset:

- containing only observations from 2013 (hint, use the filter function in the dplyr package)
- that contains only these variables: year, country, continent, health_exp_total, child_mort, life_expectancy (hint: use the select function in the dplyr package)
- contains no missing values (hint: use the drop_na() function in the tidyr package)

Your new dataset should have 178 rows and 6 variables.

Now, let's say we want to know the average and standard deviation of the child mortality in each continent.

Use the group_by() and summarise() functions of the dplyr package to calculate these.

Rounded to one decimal place, what is the average number of children that died before the age of five, per 1000 born, in Africa in 2013? (Recall that the child_mort variable gives number of deaths by age five per 1000 births.)

(Make sure you do the rounding on the accurate value, and not one already rounded by R somehow.)

```{r q4}
financing_healthcare_of_interest <- financing_healthcare |>
  filter(year == 2013) |>
  select(year, country, continent, health_exp_total, child_mort, life_expectancy) |>
  drop_na()

child_mortality_per_continent <- financing_healthcare_of_interest |>
  group_by(continent) |>
  summarize(continent=continent, child_mortality=child_mort)

child_mortality_per_continent |>
  filter(continent == "Africa") |>
  summarise(mean=child_mortality 
            |> mean() 
            |> round(1))
```

## Question 5
Now make a box and whisker plot of child mortality (y-axis / response variable) for each continent (x-axis / explanatory variable).

Hint: work with the dataset containing the 178 observations. Do not use the averages you just calculated.

Hint: Use ggplot and geom_boxplot

```{r q5}
child_mortality_per_continent |>
  ggplot(aes(x=continent, y=child_mortality)) +
  geom_boxplot() +
  theme_bw()
```

## Question 7
For a little practice with graphing, make some bivariate scatterplots of child mortality on the y-axis and total health care expenditure on the x-axis.

In one visualisation, assign colours to the points corresponding to their continent. (I.e. map the colour aesthetic to the levels in the continent variable.) Hint: use ggplot, because it will automatically make a nice key, showing which continents are which colour.

In another, make a separate graph (i.e. facet) for each continent. Hint: use ggplot because then its real easy! Hint: the ggplot default is to make the axes limits the same on all facets, but we have quite different data ranges in different continents. Using the help for facet_wrap, see if you can work out how to make the axes limits free to vary among the facets.

```{r q7}
financing_healthcare_of_interest |>
  ggplot(aes(x=health_exp_total, y=child_mort, color=continent)) +
  geom_point(size=2) +
  labs(title="Child mortality vs healthcare expenses", 
       x="Total healthcare expenses ($)",
       y="Child mortality (1000s of dead children under 5)") +
  theme(text = element_text(family = "Helvetica Neue")) +
  theme_bw()
```
```{r q7.2}
financing_healthcare_of_interest |>
  ggplot(aes(x=health_exp_total, y=child_mort)) +
  facet_wrap(~ continent, scales = "free") +
  geom_point(size=1) +
  labs(title="Child mortality vs healthcare expenses", 
       x="Total healthcare expenses ($)",
       y="Child mortality (1000s of dead children under 5)") +
  theme(text = element_text(family = "Helvetica Neue")) +
  theme_solarized() 
```
